|

### 2021-12-02 02-Chp-Systematic-Review-Tools.Rmd
End of day:

(ref:sys-rev-tools-quote) --- Bealy, 2013 [@bealy2013]

@@

<!-- TODO I might need to make it clearer thatthe tool was always intended to search both databases, as medRxiv grew out of bioRxiv categories, and no relevant records identified by the search in medRxiv -->

## Lay summary

Preprints are copies of academic manuscripts that are posted online in advance of being formally published by an academic journal. They represent an important source of scientific literature. A new software program called `medrxivr` was created to allow researchers to find preprints related to their research in a transparent and reproducible way. Development of this tool was an essential part of this thesis, as preprints represent a key source of information needed for the research reported in future chapters.

@@

Searching these preprints as part of the systematic review described in Chapter \@ref(sys-rev-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the often complex Boolean logic (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search were to go through each record, one-by-one, downloading individual citations. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

<!-- IDEA Could remake this using Diagrammr? -->

@@

Prior to development of this tool, I conducted an audit of existing tools for accessing medRxiv and bioRxiv metadata. While none address the success criteria described above, two of these tools are useful to consider to highlight the additional functionality that `medrxivr` contributes.

@@

Early versions of the tool had a reliance on scraping data directly from the repository website. Web-scraping is a fragile mechanism for extracting data, as it is entirely dependent on consistent website design and underlying code structure remaining unchanged.[@shaw2002; @laprie1992]. In the case of `medrxivr`, as the medRxiv/bioRxiv websites are regularly updated, ensuring the web-scraping performed as expected required me to regularly update or fix the script.

@@

I wrote the `medrxvir` package in R using RStudio,[@rcoreteam2019] and followed development best-practice, including development of detailed documentation, a robust unit testing framework (99% of all code lines within the package are formally tested across multiple platforms including Windows, MacOS, and Linux), and in-depth code review by two experienced, independent reviewers.

@@

<!-- TODO Clean up this section -->

<!-- REVIEW Not clear why there are two, or what are advantages or disadvantages of each. There’s a hint below (“can become unavailable during peak usage times”), but perhaps add some more -->

@@

This provides an up-to-the-minute reflection of the medRxiv preprint repository. However, this approach has two limitations. Firstly, as the API returns results as a series of pages limited to 100 records per page, downloading the entire database  <!-- TODO Get numbers here, at a given date --> requires a time-intensive process of cycling through multiple pages. Secondly, the API can become unavailable, either during peak usage times or planned maintainence windows.

To address these limitations, I provide a second method of accessing medRxiv data, called via the `mx_snapshot()` function, which allows users to access a maintained static snapshot of the database.

@@

This snapshot is created each morning at 6am using a process known as "git-scraping",[@zotero-15031] whereby the entire database is downloaded using the `mx_api_content()` function and saved as a comma separated value (CSV) file to an online server (Figure \@ref(fig:medrxivr-data-sources)). Calling `mx_snapshot()` imports this CSV into R, and has the advantage of both faster loading of the data into R (as it is imported as a single file and does not require cycling through the output of the API) and an absence of any reliance on the API.

The one limitation of this approach is that the snapshot (by its nature) will not contain details of records added to the database since it was taken. However, given that the number of records added each day is relatively low, this should pose minor issues.

@@

For example, when written in standard syntax, the search contained in the `myquery` object above would be: "((dementia **OR** alzheimer's) **AND** (lipids **OR** statins))". There is no limit to the number of search terms that can be included in each topic, nor in the number of topics that can be search for. Search terms can also contain common syntax used by systematic reviewers and health librarians, including the use of NEAR statements which allows for identification of co-localised terms, and wild-cards, which allow for alternate spellings, e.g. "randomi*s*ation" vs "randomi*z*ation".

@@

This allows users to discover which terms in their search are contributing most to the total number of results returned. This is important as part of developing a search strategy,[@bramer2018] as it allows for the key terms related to each topic to be discovered. It also aids in identifying misspelled or case-sensitive search terms, which will frequently return no results. As an example, in the search presented above, the term "alzheimer's" returns no records. This is expected, as "Alzheimer" is a proper noun and so should be capitalised, but serves to illustrate the usefulness of the reporting function.

<!-- TODO Add some text to this section about the challenges of defaults, and the fact that the autocaps option exists -->

<!-- REVIEW Can you not just make the search case insensitive? It means you often need to include capitalized and non-capitalized versions in case a word appears at the start of a sentence or is capitalized for some other reason? (I find a lot of people these days pseudo-randomly capitalize words for no good reason (like you did in “Application Programming Interface” on page 4 above!!)) -->

@@

In line with my second success criteria (Section \@ref(success-criteria)), one of the key features of the `medrxivr` is the ability for users to easily export the results of their systematic search to a reference manager. While it is a seemingly simple request, this is is one of the key ways in which `medrxivr` is set apart for other preprint search tools, including the native medRxiv/bioRxiv website search functionality.

@@

`medrxivr` alos allows users to download the full text papers for records that are deemed eligible for full-text screening (see Figure \@ref(fig:medrxivr-sr)). `mx_download()` takes the list of included records and saves the PDF for each to a folder specified by the user. This functionality is similar to the "Find Full Text" feature offered by EndNote.

@@

In addition to being used to search systematically search health-related preprint servers, as illustrated in the systematic review presented in Chapter \@ref(sys-rev-heading), `medrxivr` has other uses.

For example, I led a descriptive analysis of the change in data availability statements between preprinted and published versions of the same manuscript, stratified by journal data sharing policy access, underpinned by preprint meta-data provided by `medrixvr`. By comparing the preprinted and published versions of the data availability statement, I could examine the same manuscript (same content, authors and funders) under two different publication policies, and examine whether stricter policies which require data sharing as a condition of publication actually result in increased data availability. We found some evidence that data availability statements more frequently described open data on publication when the journal mandated data sharing compared to when the journal did not mandate data sharing. This study has since been published in PLOS One, and a copy is included in Appendix \@ref(published-papers).

@@

published <- med_res$n[2]
publication_n <- sum(med_res$n)
publish <- paste0(num_to_text(published,T)," (", round(published/publication_n*100,1),"%)")
```

<!-- TODO Need to finish this bit and move it to the use-case section! -->

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-sys-rev-tools)). `r publish` of the `r publication_n` records posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting,[@abdill2019popularity] indicating that a non-insignificant number of preprints are never formally published but remain accessible as preprints.

In short, these use cases illustrate that easy access to medRxiv/bioRxiv metadata has applications beyond systematic searching of preprints by enabling meta-research/methodological analyses.

@@

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, meaning some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

There is also the potential that the cross-section of literature posted on medrxiv/bioRxiv is substantially different those suffering from publication bias (studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant).[@song2010] This is because simply lowering the barriers to publication may well encourage authors to published "null" results, but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[@rosenthal1979]

<!-- TODO Clean-up this section -->

<!-- REVIEW I don’t think anyone would assume that preprints would help much here… I’d have thought the main argument is getting hold of evidence in a more timely manner.-->

<!-- TODO replicate this analysis -->

It is likely too early (and likely too methodologically difficult) to tell whether the increased popularity and acceptance of preprint repositories will have any effect of the availability of research that was not considered "publishable" at other venues.

<!-- TODO EMBASE will index medRxiv/bioRxiv from Christmas - need to make a note of this somewhere -->

@@

Part of the motivation for creating the `medrxivr` tool was a belief that the development and distribution of open source scripts and tools should be a fundamental part of evidence synthesis research.[@goldacre2019; @mckiernan2016] In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts that have a similar, or related, functionality - in fact, following development of the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package, which allows for importing medRxiv metadata into R but does not provide search functionality).[@fraser2020rbiorixv] If these scripts continue to be developed in private and are never shared or publicised, this will inevitably hamper the efforts of evidence synthesis community, not only in terms of duplication of time and effort but also due to lost opportunities for collaboration.[@mckiernan2016] Creating and sharing well-documented packages, the recognised standard for sharing code in R, represents one way to reduce this inefficiency.[@vuorre2020]

@@

-   I have outlined the motivation for developing this tool in relation to this thesis - more specifically, that it was used to perform systematic and reproducible searches of a key literature sources used in the comprehensive systematic review described in Chapter \@ref(sys-rev-heading).

>>>>

(ref:sys-rev-tools-quote) --- Chris Beely, 2013 [@beely2013]

@@

::: {.laybox data-latex=""}

## Lay summary {-}

Preprints are copies of academic manuscripts that are posted online in advance of being formally published by an academic journal. They represent an important source of scientific literature. A new software program called `medrxivr` was created as part of this thesis to allow researchers to find preprints related to their research in a transparent and reproducible way. Development of this tool was an essential part of this thesis, as preprints represent a key source of information needed for the research reported in future chapters.

:::

@@

Searching these preprints as part of the systematic review described in Chapter \@ref(sys-rev-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

@@

Prior to development of this tool, I conducted an audit of existing tools for accessing medRxiv and bioRxiv metadata. While none addresses the success criteria described above, two of these tools are useful to consider to highlight the additional functionality that `medrxivr` contributes.

@@

Early versions of the tool had a reliance on scraping data directly from the repository website. Web-scraping is a fragile mechanism for extracting data, as it is entirely dependent on consistent website design and underlying code structure remaining unchanged.[@shaw2002; @laprie1992]. In the case of `medrxivr`, as the medRxiv/bioRxiv websites are regularly updated, ensuring the web-scraping performed as expected required that I regularly update or fix the script.

@@

I wrote the `medrxvir` package in R using RStudio,[@rcoreteam2019] and followed development best-practice, including detailed documentation, a robust unit testing framework (99% of all code lines within the package are formally tested across multiple platforms including Windows, MacOS, and Linux), and in-depth code review by two experienced, independent reviewers.

@@

@@

This provides an up-to-the-minute reflection of the medRxiv preprint repository. However, this approach has two limitations. Firstly, as the API returns results as a series of pages limited to 100 records per page, downloading the entire database requires a time-intensive process of cycling through each page. Secondly, the API can become unavailable, either during peak usage times or planned maintainence windows. To address these limitations, I provide a second method of accessing medRxiv data, called via the `mx_snapshot()` function, which allows users to access a maintained static snapshot of the database.

@@

This snapshot is created each morning at 6am using a process known as "git-scraping",[@zotero-15031] whereby the entire database is downloaded using the `mx_api_content()` function and saved as a comma separated value (CSV) file in an online repository (Figure \@ref(fig:medrxivr-data-sources)). Calling `mx_snapshot()` imports this CSV into R, and has the advantage of both faster loading of the data into R (as it is imported as a single file and does not require cycling through the output of the API) and an absence of any reliance on the API. The one limitation of this approach is that the snapshot (by its nature) will not contain details of records added to the database since it was taken. However, given that the number of records added each day is relatively low, this should pose minor issues.

@@

For example, when written in standard syntax, the search contained in the `myquery` object above would be: "((dementia **OR** alzheimer's) **AND** (lipids **OR** statins))". There is no limit to the number of search terms that can be included in each topic, nor in the number of topics that can be searched for. Search terms can also contain common syntax used by systematic reviewers and health librarians, including the NEAR operator for the identification of co-localised terms and wild-cards which allow for alternate spellings (e.g. "randomi*s*ation" vs "randomi*z*ation").

@@

This allows users to discover which terms in their search are contributing most to the total number of results returned. This is important as part of developing a search strategy,[@bramer2018] as it allows for the key terms related to each topic to be discovered. It also aids in identifying misspelled or case-sensitive search terms, which will frequently return no results. As an example, in the search presented above, the term "alzheimer's" returns no records. This is expected, as "Alzheimer's" is a proper noun and so should be capitalised, but serves to illustrate the usefulness of the reporting function.

In hindsight, the decision to make the search case-sensitive was a poor one as users reported being unaware of this setting when searchin. However, changing the default behaviour of `mx_search()` to be case-insensitive is problematic, as it could result in different numbers of hits based solely on which version of the package the user has installed. To address this, `mx_search()` now contains an `autocaps` argument, which when set to `TRUE` will automatically search for both capitalised and uncapitalised versions of the search term.

@@

In line with my second success criteria (Section \@ref(success-criteria)), one of the key features of the `medrxivr` is the ability for users to export easily the results of their systematic search to a reference manager. While it is a seemingly simple request, this is is one of the key ways in which `medrxivr` is set apart for other preprint search tools, including the native medRxiv/bioRxiv website search functionality.

@@

`medrxivr` also allows users to download the full text papers for records that are deemed eligible for full-text screening (see Figure \@ref(fig:medrxivr-sr)). `mx_download()` takes the list of included records and saves the PDF for each to a folder specified by the user. This functionality is similar to the "Find Full Text" feature offered by EndNote.

@@

In addition to being used to search systematically health-related preprint servers, as illustrated in the systematic review presented in Chapter \@ref(sys-rev-heading), `medrxivr` has other uses.

For example, I led a descriptive analysis of the change in data availability statements between preprinted and published versions of the same manuscript, stratified by journal data sharing policy access, underpinned by preprint meta-data provided by `medrixvr`. By comparing the preprinted and published versions of the data availability statement, I could examine the same manuscript - same content, authors and funders - under two different publication policies (preprint server vs peer-review journal). This comparison provides evidence on whether stricter policies which require data sharing as a condition of publication result in increased data availability. The analysis found some evidence that data availability statements more frequently described open data on publication when the journal mandated data sharing, compared to when the journal did not mandate data sharing. This study has since been published,[@mcguinness2020DAScomparison] and a copy is included in Appendix \@ref(published-papers).

@@

published <- med_res$n[2]
publication_n <- sum(med_res$n)
publish_percent <- paste0("(", round(published/publication_n*100,1),"%)")
```

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-sys-rev-tools)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a non-insignificant number of preprints are never formally published but remain accessible as preprints.

In summary, the meta-research/methodological analyses described above illustrate that easy access to medRxiv/bioRxiv metadata has applications beyond the inclusion of preprints in systeamtic review.

@@

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, and so some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

There is also the potential that the cross-section of literature posted on medrxiv/bioRxiv is substantially different to those suffering from publication bias (studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant).[@song2010] This is because simply lowering the barriers to publication may well encourage authors to published "null" results, but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[@rosenthal1979]

It is probably too early (and likely too methodologically difficult) to tell whether the increased popularity and acceptance of preprint repositories will have any effect of the availability of research that was not considered "publishable" at other venues.

@@

Part of the motivation for creating the `medrxivr` tool was a belief that the development and distribution of open source scripts and tools should be a fundamental part of evidence synthesis research.[@goldacre2019; @mckiernan2016] In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts that have a similar, or related, functionality - in fact, following development of the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package, which allows for importing medRxiv metadata into R but does not provide search functionality).[@fraser2020rbiorixv] If these scripts continue to be developed in private and are never shared or publicised, this will inevitably hamper the efforts of the evidence synthesis community, not only in terms of duplication of time and effort but also due to lost opportunities for collaboration.[@mckiernan2016] Creating and sharing well-documented packages, the recognised standard for sharing code in R, represents one way to reduce this inefficiency.[@vuorre2020]

@@

-   I have outlined the motivation for developing this tool in relation to this thesis - more specifically, that it was used to perform systematic and reproducible searches of key literature sources used in the comprehensive systematic review described in Chapter \@ref(sys-rev-heading).

### 2021-12-02 04-Chp-Systematic-Review-Results.Rmd
End of day:

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzheimer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published and extracted results including SNPs in the APoE4 genetic region (see the following section for a discussion of the bias this introduces).

@@

One item of particular interest is the attenuation of any effects identified by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As discussed in the introduction (see Section \@ref(intro-basic-science)), an increasing number of ApoE4 alleles is a major independent risk factor for Alzheimer's disease and so violates the exclusion restriction criteria of Mendelian randomisation studies (Figure \@ref(fig:mrAssumptions)). In all cases, excluding variants in the ApoE region attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where ApoE variants were not sufficient identified and excluded, leading the published paper to report a protective effect of LDL-c on Alzheimer's (`r estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed to exclude a larger area around ApoE which attenuated the finding to the null.[@benn2017comment]

>>>>

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzheimer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published and extracted results including SNPs in the _ApoE4_ genetic region (see the following section for a discussion of the bias this introduces).

@@

One item of particular interest is the attenuation of any effects identified by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the _ApoE4_ gene region. As discussed in the introduction (see Section \@ref(intro-basic-science)), an increasing number of _ApoE4_ alleles is a major independent risk factor for Alzheimer's disease and so violates the exclusion restriction criteria of Mendelian randomisation studies (Figure \@ref(fig:mrAssumptions)). In all cases, excluding variants in the _ApoE_ region attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where _ApoE_ variants were not sufficient identified and excluded, leading the published paper to report a protective effect of LDL-c on Alzheimer's (`r estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed to exclude a larger area around _ApoE_ which attenuated the finding to the null.[@benn2017comment]

### 2021-12-02 05-Chp-CPRD-Analysis.Rmd
End of day:

This chapter represents an extended version of a preprint manuscript, a copy of which is available in Appendix \@ref(published-papers).

@@

A further limitation stems from the possibility of uncontrolled confounding due to genetic factors. The number of ApoE $\epsilon4$ alleles represents the strongest genetic risk factor for Alzheimer's disease, but also substantially increases LDL cholesterol levels,[@bennet2007] potentially prompting treatment with a statin or other lipid regulating agent. I was unable to control for ApoE genotype in this analysis as I did not have access to genetic data on participants. As a result, any protective association between LRA use and the Alzheimer's disease outcomes may be masked by residual negative confounding by ApoE.

>>>>

This chapter represents an extended version of a preprinted manuscript, a copy of which is available in Appendix \@ref(published-papers).

@@

A further limitation stems from the possibility of uncontrolled confounding due to genetic factors. The number of _ApoE_ $\epsilon4$ alleles represents the strongest genetic risk factor for Alzheimer's disease, but also substantially increases LDL cholesterol levels,[@bennet2007] potentially prompting treatment with a statin or other lipid regulating agent. I was unable to control for _ApoE_ genotype in this analysis as I did not have access to genetic data on participants. As a result, any protective association between LRA use and the Alzheimer's disease outcomes may be masked by residual negative confounding by ApoE.
